Smart Representations: {R}ationality and Evolution in a Richer Environment

### 2015-12-14 paper/paper.tex
Paolo edits with conclusion

\title{Smart Representations: {R}ationality, Regret, Evolution \& the Environment}
\author{Paolo Galeazzi \& Michael Franke} \date{current version: November 5, 2015}

@@

\section{Introduction / Motivation}
\label{sec:intr--motiv}

Evolutionary game theory has become an established philosophical tool for probing into
conceptual issues whose complexity requires mathematical modelling. However, most applications
look at evolutionary selection of behavior in a single \emph{stage game}. There are two common
simplifications in this approach. Firstly, the stage game is usually considered fixed, closed
and immutable. Secondly, what evolves are behavioral strategies for the stage game. It has
consequently been argued that, instead of focusing on a fixed, immutable stage game and the
behavioral selection for that stage game alone, we should rather study the evolutionary
competition of \emph{choice mechanisms} under selective pressure in a rich and variable
environment \citep[e.g.][]{FawcettHamblin2013:Exposing-the-be,McNamara2013:Towards-a-Riche}. 
But only little attention has been paid to modeling the evolutionary competition between choice
mechanisms in a game-theoretical framework \citep[but
see][]{ZollmanSmead2010:Plasticity-and-,SmeadZollman2013:The-Stability-o}.\todo{More references
  Harley 1981!!!, O Conor \dots}

If agents deal with a rich and varied environment, they face many different choice
situations. A crucial part of any choice mechanism is a \emph{subjective representation} of the
choice situation, e.g., the agent's manner of representing their preferences and her way of
forming decision relevant beliefs in the case of uncertainty. Since almost any reasonable
learning or decision-making process will make use of some (subjective, agent-internal (possibly
unconscious)) representation of the quality of choice options, the question of which such
representations are ecologically valuable (i.e., lead to high fitness) is central and has been
subject of recent interest \citep[e.g.][]{algweib13,DekElyYlan07,Sam01,RobSam11}. Na\"ive intuition might convince us that preferences that veridically
represent the actual, true fitness measure that evolutionary selection is based upon must
surely be the best. We demonstrate that, if we take into account how agents represent
uncertainty, this na\"ive conclusion is premature. Under certain reasonable epistemic
conditions, non-veridical preference representations can be evolutionarily beneficial.

To show this, we study the evolutionary competition of choice mechanisms, or more specifically
the competition of subjective representations of choice situations, in a variable
environment. Concretely, we look at evolutionary ``meta-games'' whose ``acts'' are types of
subjective representations of games and whose ``payoffs'' represent the average expected
fitness that choice mechanisms using these representation types would accrue when playing
arbitrary games (from a given class, with a given occurrence probability). In this sense, a
meta-game captures (the modeller's) assumptions about the relevant statistics of the
environment in which evolutionary forces operate, while we are still able to use standard
methods from evolutionary game theory, like stability or evolutionary dynamics.

\todo[inline]{add overview}

@@

\begin{definition}[Rational choice from maximization of expected utility]
  \label{def:rationality}
  Let $S$ be a (finite) set of states of the world and $A$ a (finite) set of actions. Given a
  probability distribution $P$ over $S$ and a utility function
  $u \mycolon S \times A \rightarrow \mathbb{R} $, an action $a^* \in A$ is a rational choice
  only if it maximizes the \emph{(subjective) expected utility}
  $\EU(a) = \sum_{s \in S} P(s) \ u(s, a)$.
\end{definition}

@@

If beliefs and preferences are subjective, there is room for \emph{rationalization} or
\emph{redescriptionism} of observable behavior. Prospect theory is, perhaps, the most prominent
example. \citet{KahnemannTversky1979:Prospect-Theory} looked at experimental data from human
choices in suitably defined tasks and demonstrated that what appears to be violations of
rationality norms can be explained on the assumption that subjects' beliefs and preferences
deviate systematically from the objectively given parameters in the presented choice task
(e.g., by taking into account reference points and diminishing returns for both gains and
losses).  Similarly, tinkering with the notion of subjective utility, in particular by
including social preferences such as fairness, reciprocity or similar, allows us to describe as
rational empirically observed behavior in situations of social decision making, such as the
prisoners' dilemma or public goods games, that might otherwise appear irrational \citep[e.g.][]{fehrschmidt99,charrab02}.

@@

An alternative to descriptive vs.~normative views on rationality is \emph{adaptationism}
\citep[e.g.][]{HagenChater2012:Decision-Making}. \todo{more references} Adaptationism aims to
explain choice behavior by appeal to evolutionary considerations. The key idea is that decision
makers (human and other life forms) have evolved choice mechanisms which were fitness enhancing
in the relevant domains of choice making, at the relevant time of evolution. When it comes to
preferences, this is clearly related to Alchian and Friedman's argument that firms
\emph{should} optimize that which makes them ``live,'' so to speak. For biological agents, the
argument would be that subjective preferences \emph{should} stay as close as possible to whatever determines objective
fitness.

We would like to put this latter idea to the test. We present a formal argument intended to
support the conclusion that subjective preferences that differ from objective fitness can be
adaptive. We argue that it is important to take into account the potential mutual dependence of
subjective preferences and subjective beliefs in adaptationist approaches to decision
making. In other words, we focus on the evolutionary competition between \emph{subjective
  representation schemes}: general ways of forming subjective preferences and
subjective beliefs. We show that subjective, non-veridical preference representations based on
a notion of regret \citep[e.g.][]{Savage1951:The-theory-of-s,}
can be adaptive and even outperform objective, veridical representations of the actual, true
fitness, when agents, at least on occasion, hold \emph{imprecise probabilistic beliefs}
(\textcolor{red}{CITATIONS}) and make decisions based on a security choice rule (here: the
maximin rule applied to expected utility). The next section introduces all of these notions
(precise/imprecise beliefs, veridical/regret representations, \dots) in detail.

@@

For the purposes of this paper, we think of a choice rule as a function that takes a
(strategic) game as input and outputs a choice of action for that game. The literature has
suggested a number of choice rules that aspire to capture a normative element of rational
choice. We focus here on three classical choice rules: (i) maximin, (ii) Laplace and (iii)
regret minimization.

@@

Perhaps the most well-known choice rule that does not require explicitly stating the DM's
beliefs is the \emph{maximin rule}. It tells our agent to choose act $\act^*$ only if it
maximizes payoffs for all worst case scenarios. For a strategic game with 2 players this can be
formalized as
$\act^* \in \argmax_{\myact{i} \in \Acts_i} \min_{\myact{j} \in \Acts_j} \pi_i(\myact{i},
\myact{j})$.
For the example in Figure~\ref{coordgame1}, the maximim rule looks at the minima in each row,
asking: what is the worst possible outcome for each row act? It is $0$ in either
case. Consequently, the maximin rule would not favor either act: players choosing by the
maximin rule would be indifferent between acts $I$ and $II$.

@@

\begin{definition}[Regret, regret minimization] \label{defn:regret} Given a strategic game
  $ G=\langle N, (A_i , \pi_i)_{i \in N} \rangle $, the \emph{pairwise regret} of player $i$'s
  act $a_i$ against profile $a_{-1}$ is
  $\text{reg}(a_i,a_{-i}):= \pi_i(a_i^\$,a_{-i})-\pi_i(a_i,a_{-i}) $, where
  $a_i^\$ \in \argmax_{a_i \in A_i} \pi_i(a_i, a_{-i})$ denotes $i$'s best reply to the
  opponents' actions $a_{-i}$. The \emph{regret} of $a_i$ is then
  $\text{reg}(a_i):= \text{max}_{a_{-i}\in A_{-i}} \pi_i(a_i^\$,a_{-i})-\pi_i(a_i,a_{-i}) $. An
  action $a^{*}_i $ is regret minimizing if
  $a^{*}_i \in \text{min}_{a_i} \text{max}_{a_{-i}} \pi_i(a_i^\$,a_{-i})-\pi_i(a_i,a_{-i}) $.
\end{definition}

@@

% \begin{itemize}
% \item \textcolor{red}{add here:}
%   \begin{itemize}  
%   \item outcome of rationalizability (max-EU)
%   \item outcome of maximin rule
%   \item outcome of Laplace rule
%   \end{itemize}

@@

Regret minimization is intimately related to the maximin rule. Let the \emph{negative regret
  transformation} of a strategic game $G$ be the game $G'$ derived from $G$ by replacing the payoff
function $\pi$ of $G$ with $\pi'$ such that $\pi'_i(a_i, a_{-i}) = - reg(a_i,a_{-1})$, i.e.,
the payoff of all players is the pairwise negative regret (under the payoffs from $G$). For
example, the negative regret transform of the game in Figure~\ref{coordgame1} would have the
following payoff matrix, i.e., the negatives of the pairwise regrets given in
Figure~\ref{coordgame1reg}:

@@

\noindent Acts selected by the maximin rule for the negative regret transformation of $G$ are
always exactly the acts selected by regret minimization for the original game $G$.\todo{can we
  cite a paper that proves this? should we give a proof ourselves?}

This means that we can think of the maximin rule and regret minimization as essentially the
same operation on different subjective representations of preference: while maximin considers
actual objective payoffs (as defined by the modeller's game), regret minimization considers
non-veridical, regret-based subjective preferences. As the example of Figure~\ref{coordgame1}
shows, different subjective preferences can give rise to different choice prescriptions all
else equal. While maximin is indifferent between $I$ and $II$, regret minimization uniquely
selects $II$.

In sum, a choice rule is a normative prescription of action choices for a rational agent in any
game situation. Choice rules may rest on implicit assumptions about what the DM believes (e.g.,
when interpreting Laplace rule as rational choice under flat beliefs). They may also make
implicit assumptions about what the DM prefers (e.g., when reconstructing regret minimization
in terms of maximin-choices on the regret transformation). This is rather messy. To disentangle
and avoid confusion, we turn to choice mechanisms.

@@

A \emph{choice mechanism} is a map from choice situations to action choices that, unlike choice
rules, contains an explicit level of subjective representation of a decision situation. For our
purposes, we should think of a choice mechanism as a pair of functions $\gamma$ and $\kappa$,
where $\gamma$ is a \emph{subjective representation scheme} and $\kappa$ is an \emph{action
  selection function} (see FigureXYZ).\todo{make figure when notation and terminology are
  fixed}\ The subjective representation scheme $\gamma$ maps a game (or more generally any
decision situation) onto a subjective representation of the game (or choice situation) and
other decision-relevant beliefs (e.g., about what the opponent is likely to play). The action
selection function $\kappa$ takes such a subjective representation that $\gamma$ returns and
maps it onto a concrete action choice. The crucial bit for our purposes here is the
representation function $\gamma$, because we are interested in studying which ways of
conceptualizing a decision situation are evolutionarily successful.

A choice mechanism could be a set of distinct heuristics (the DM's \emph{adaptive toolbox})
that have little in common
\citep[e.g.][]{TverskyKahnemann1981:The-Framing-of-,GigerenzerGoldstein1996:Reasoning-the-F,ScheibehenneRieskamp2013:Testing-the-Ada}. But
to keep matters simple and implementable for a start, we will here only look at choice
mechanisms that differ only with respect to their subjective representation function $\gamma$,
which in turn is conceived of as a \emph{uniform} way of forming preference and belief
\emph{types}: we look at general \emph{subjective representation schemes}.

Formally, given a class $ \mathcal{G} $ of strategic games,
$\gamma_i: \mathcal{G} \rightarrow \Theta_i \times E_i$ is a function that assigns a subjective
preference $\theta_i \in \Theta_i$ and an epistemic state $e_i \in E_i$ to any game
$G \in \mathcal{G}$ for some agent $i$. A subjective preference $\theta_i$ for game $G$ and
agent $i$ is just a payoff function for $i$ suitable to the acts and co-players in
$G$. Crucially, this subjective payoff function can be different from the actual payoff function $\pi^G$ of $G$,
which we interpret as the objective evolutionary fitness. An epistemic state $e_i$ for game $G$ and agent
$i$ is some specification of $i$'s uncertainty about the co-players' behavior. Finally, an
action selection function for agent $i$ is a function
$\kappa_i: \Theta_i \times E_i \rightarrow \pow{A_i} $, that associates any subjective
representation of the game $ (\theta_i,e_i) \in \Theta_i \times E_i$ with a set of prescribed
actions for $i$ in $G$ under $i$'s subjective representation of $G$. We drop the index $i$ for
readability in the following.

We focus on two (\textit{subjective}) \emph{preference types} and two \emph{belief types}. We
view a preference type as a function $\tau: \mathcal{G} \rightarrow \Theta$, and a belief type
as a general disposition to form beliefs about the co-player’s behavior. The two possible
dispositions that we take account of are precise probabilistic belief and imprecise belief, as
we will explain in the next section. Since we are most interested here in studying whether
evolutionary pressure could support non-veridical subjective preferences, the two preference
types that we consider are: the veridical type $\tau^\pi$, who uses the actual payoff function
$\pi$ of the actual game $G$, and the regret type $\tau^{reg}$, who looks at the regret
transformation of the actual game $G$. Finally, the action selection function $\kappa$ that we
use in this paper is \textit{maxmin expected utility} [GS89], for reasons that will be clear
and justified in the following two sections.

@@

% The following subsection describes what these belief types are and how a single action
% selection function (maximin expected utility) gives rise to an intersting parallel with
% classical choice rules.

@@

Consider the following problem. I have, on the desk in front of me, a bag containing either red or black marbles. There is no further information about the distribution of the two colors. What is then the probability that I will draw a red marble? As in Walley91-96, we are not talking about physical probability, but rather about epistemic probability, i.e., the subjective probability that an agent attributes to an event.
In the literature there are two traditional ways of answering this kind of questions. According to standard Bayesianism, one should always be able to determine a precise betting rate (that immediately translates into a precise probability distribution), which is fair, in the sense that the DM is indifferent between betting on or against red at the fair rate, and would bet on red at any rate that is more favourable than the fair rate and against red at any rate that is less favourable. In the case of the bag on my desk, a strict Bayesian, in the absence of any further information, would probably rely on the \textit{principle of insufficient reason} and answer that the DM should have a uniform distribution that considers all the outcomes equipossible. \\
The second answer involves imprecise probabilities and is related to the literature about unmeasurable uncertainty. Many authors argued in favor of this kind of approach \citep[e.g.][]{levi74,gardsah82,walley96}. Since it is consistent with the information available that all the marbles are red, or that no marble is red, the resulting model should take it into account and specify a lower probability of red $\underline{P}(R)$ and an upper probability of red $\overline{P}(R)$ that define an interval of possible probabilistic beliefs [$\underline{P}(R), \overline{P}(R)$]. Given the information available in this case we would have $\underline{P}(R)=0$ and $\overline{P}(R)=1$. This approach appears particularly relevant in a game theoretical context. Indeed, in a recent paper \citet{BattCerrMM15} write

@@

In evolutionary game theory, for instance, players in a population obviously face uncertainty about the composition of the population that they are part of, and consequently about the co-player that they are randomly paired with at each round and about the co-players' action. We represent this uncertainty in two different ways that correspond to the two approaches discussed above: a Bayesian uniform distribution and an imprecise probability set. This amounts to having two different belief types, $ \overline{\mu}$ and $ \Delta(A_{-i})$, where $\overline{\mu}$ stands for the uniform distribution and $\Delta(A_{-i})$ is the probability set, defined as follows:

@@

\item $ \Delta(A_{-i}):=\lbrace \mu \in \mathbb{R}^{A_{-i}}: \sum_{a_{-i} \in A_{-i}} \mu(a_{-i}) = 1 \rbrace$;

\item $\overline{\mu} \in \Delta(A_{-i})$ is the probability measure over the co-players' actions such that: $\forall a_{-i} \in A_{-i}$, $\overline{\mu}(a_{-i})= \frac{1}{|A_{-i}|}$.

@@

Hence, we consider two possible epistemic states in the face of uncertainty. The belief type $\Delta(A_{-i})$ represents a situation of non-Bayesian radical uncertainty, where the players are not able to narrow down the set of probability distributions to a single one and they consider all of them as possible. The other belief type $\overline{\mu}$ conceptualizes the uncertainty about the population by using the principle of insufficient reason and ascribes equal probability to all the possible alternatives. As we have seen, both options are reasonable and justifiable ways to deal with situations of uncertainty. \\

In decision theory, the imprecise probability model is in line with most of the representation results of decision making under uncertainty \citep[e.g.][]{gilsch89,KlibMarMuk05}, and seems justified by empirical observations too.

@@

\begin{fact} \label{fact:singleton probability set}
Whenever the probability set is a singleton, maximinization of
expected utility coincides with maximization of expected utility.
\end{fact}

@@

By fixing the action selection component to maximin expected utility, and allowing two belief and
two preference types, we regain all the choice rules considered in sectionNN from a more general and unified standpoint, as it is summed up in
Table~\ref{tab:CMs}. \\
We are now in the position of formally deriving behavior from a pair $(\theta, e) \in \Theta \times E$.

@@

It is then worth stressing that $\kappa(\tau^{\pi}, \Delta(A_{-i}))$ turns out to be the actual payoff
maximinimizer, $\kappa(\tau^{reg}, \Delta(A_{-i}))$ corresponds to the regret minimizer,
$\kappa(\tau^{\pi}, \overline{\mu})$ is the expected actual payoff maximizer (with flat beliefs), and
$\kappa(\tau^{reg}, \overline{\mu})$ is the expected regret minimizer (with flat beliefs).
Moreover, different subjective preferences and belief types can give rise to identical action predictions,
either in a restricted class of games, or in general. This is also the case here.

>>>>

\title{Smart Representations: {R}ationality and Evolution in a Richer  Environment}
\author{Paolo Galeazzi \& Michael Franke} \date{current version: November 5, 2015}

@@

\section{Introduction}
\label{sec:intr--motiv}

\begin{quotation} \textit{This is what I aim at, because the point of philosophy is to start with something so simple as not to seem worth stating, and to end with something so paradoxical that no one will believe it.} \citep{Russ18} \end{quotation}

Evolutionary game theory has become an established philosophical tool for probing into
conceptual issues whose complexity requires mathematical modelling. However, most applications
look at evolutionary selection of behavior in a single \emph{stage game}. There are two common
simplifications in this approach. Firstly, the stage game is usually considered fixed, closed
and immutable. Secondly, what evolves are behavioral strategies for the stage game. It has
consequently been argued that, instead of focusing on a fixed, immutable stage game and the
behavioral selection for that stage game alone, we should rather study the evolutionary
competition of \emph{choice mechanisms} under selective pressure in a rich and variable
environment \citep[e.g.,][]{FawcettHamblin2013:Exposing-the-be,McNamara2013:Towards-a-Riche}. 
But only little attention has been paid to modeling the evolutionary competition between choice
mechanisms in a game-theoretical framework \citep[but
see][]{ZollmanSmead2010:Plasticity-and-,SmeadZollman2013:The-Stability-o}.\todo{More references
  Harley 1981!!!, O Conor \dots}

If agents deal with a rich and varied environment, they face many different choice
situations. A crucial part of any choice mechanism is a \emph{subjective representation} of the
choice situation, e.g., the agent's manner of representing their preferences and her way of
forming decision relevant beliefs in the case of uncertainty. Since almost any reasonable
learning or decision making process will make use of some (subjective, agent-internal, possibly
unconscious) representation of both beliefs and the quality of choice options, the question of which such
representations are ecologically valuable (i.e., lead to high fitness) is central and has been
subject of recent interest also in economics literature \citep[e.g.,][]{algweib13,DekElyYlan07,RobSam11}. Na\"ive intuition might convince us that preferences that veridically
represent the actual, true fitness measure that evolutionary selection is based upon must
surely be the best. We demonstrate that, if we take into account the different ways agents may cope with
uncertainty, this na\"ive conclusion is premature. Under certain reasonable epistemic
conditions, non-veridical preference representations can be evolutionarily beneficial.

To show this, we study the evolutionary competition of choice mechanisms, or more specifically
the competition of subjective representations of choice situations, in a variable
environment. Concretely, we look at evolutionary ``meta-games'' whose ``acts'' are choice mechanisms and whose ``payoffs'' represent the average expected
fitness that different choice mechanisms would accrue when playing
arbitrary games (from a given class, with a given occurrence probability). In this sense, a
meta-game captures (the modeller's) assumptions about the relevant statistics of the
environment in which evolutionary forces operate, while we are still able to use standard
methods from evolutionary game theory, like stability or evolutionary dynamics.

Section \ref{sec:rati--subj} overviews different accounts of rationality in decision making and relates the approach we take here to the existing literature on the topic. Section \ref{sec:basic-notions} introduces the various decision- and game-theoretic notions that will be later used in the evolutionary analysis. Sections \ref{sec:model} and \ref{sec:basic-model-1} define the evolutionary framework, and Section \ref{sec:results:-basic-model} presents the main results of the paper. Section \ref{sec:extensions} extends the evolutionary analysis by introducing more choice mechanisms in the model and by investigating the evolutionary performances of different choice mechanisms when facing arbitrary decision problems. Finally, in Section \ref{sec:conclusion} we comment on the results achieved and future perspectives.

@@

\begin{definition}[Rational choice from maximization of expected utility]
  \label{def:rationality}
  Let $S$ be a (finite) set of states of the world and $A$ a (finite) set of actions. Given a
  probability distribution $\mu$ over $S$ and a utility function
  $u \mycolon S \times A \rightarrow \mathbb{R} $, an action $a^* \in A$ is a rational choice
 if it maximizes the \emph{(subjective) expected utility} 
  $\EU(a) = \sum_{s \in S} \mu(s) \ u(s, a)$.
\end{definition}

@@

If beliefs and preferences are subjective, there is room for \emph{rationalization} or
\emph{redescriptionism} of observable behavior. Prospect theory is, perhaps, the most prominent
example. \citet{KahnemannTversky1979:Prospect-Theory} looked at experimental data from human
choices in suitably defined tasks and demonstrated that what appears to be violations of
rationality norms can be explained on the assumption that subjects' beliefs and preferences
deviate systematically from the objectively given parameters in the presented choice task
(e.g., by taking into account reference points and diminishing returns for both gains and
losses).  Similarly, tinkering with the notion of subjective utility, in particular by
including social preferences such as fairness, reciprocity or similar, allows us to describe as
rational empirically observed behavior in situations of social decision making, such as the
prisoners' dilemma or public goods games, that might otherwise appear irrational \citep[e.g.,][]{fehrschmidt99,charrab02}.

@@

An alternative to descriptive vs.~normative views on rationality is \emph{adaptationism}
\citep[e.g.,][]{HagenChater2012:Decision-Making}. \todo{more references} Adaptationism aims to
explain choice behavior by appeal to evolutionary considerations. The key idea is that decision
makers (human and other life forms) have evolved choice mechanisms which were fitness enhancing
in the relevant domains of choice making, at the relevant time of evolution. When it comes to
preferences, this is clearly related to Alchian and Friedman's argument that firms
\emph{should} optimize that which makes them ``live,'' so to speak. For biological agents, the
argument would be that subjective preferences \emph{should} stay as close as possible to whatever determines objective
fitness.

We would like to put this latter idea to the test. We present a formal argument intended to
support the conclusion that subjective preferences that differ from objective fitness can be
adaptive. We argue that it is important to take into account the potential mutual dependence of
subjective preferences and subjective beliefs in adaptationist approaches to decision
making. In other words, we focus on the evolutionary competition between \emph{subjective
  representation schemes}: general ways of forming subjective preferences and
subjective beliefs. We show that subjective, non-veridical preference representations based on
a notion of regret \citep[e.g.,][]{Savage1951:The-theory-of-s,}
can be adaptive and even outperform objective, veridical representations of the actual, true
fitness, when agents, at least on occasion, hold \emph{imprecise probabilistic beliefs}
\citep[e.g.,][]{gilsch89,levi74,gardsah82} and make decisions based on a security choice rule (here: the
maxmin rule applied to expected utility). The next section introduces all these notions
(precise/imprecise beliefs, veridical/regret representations, \dots) in detail.

@@

For the purposes of this paper, we think of a choice rule as a function that takes a
(strategic) game as input and outputs a choice of action for that game. The literature has
suggested a number of choice rules that aspire to capture a normative element of rational
choice. We focus here on three classical choice rules: (i) maxmin, (ii) Laplace and (iii)
regret minimization.

@@

Perhaps the most well-known choice rule that does not require explicitly stating the DM's
beliefs is the \emph{maxmin rule}. It tells our agent to choose act $\act^*$ only if it
maximizes payoffs for all worst case scenarios. For a strategic game with 2 players this can be
formalized as
$\act^* \in \argmax_{\myact{i} \in \Acts_i} \min_{\myact{j} \in \Acts_j} \pi_i(\myact{i},
\myact{j})$.
For the example in Figure~\ref{coordgame1}, the maximim rule looks at the minima in each row,
asking: what is the worst possible outcome for each row act? It is $0$ in either
case. Consequently, the maxmin rule would not favor either act: players choosing by the
maxmin rule would be indifferent between acts $I$ and $II$.

@@

\begin{definition}[Regret, Regret Minimization] \label{defn:regret} Given a strategic game
  $ G=\langle N, (A_i , \pi_i)_{i \in N} \rangle $, the \emph{pairwise regret} of player $i$'s
  act $a_i$ against profile $a_{-1}$ is
  $\text{reg}(a_i,a_{-i}):= \pi_i(a_i^\$,a_{-i})-\pi_i(a_i,a_{-i}) $, where
  $a_i^\$ \in \argmax_{a_i \in A_i} \pi_i(a_i, a_{-i})$ denotes $i$'s best reply to the
  opponents' actions $a_{-i}$. The \emph{regret} of $a_i$ is then
  $\text{reg}(a_i):= \text{max}_{a_{-i}\in A_{-i}} \pi_i(a_i^\$,a_{-i})-\pi_i(a_i,a_{-i}) $. An
  action $a^{*}_i $ is regret minimizing if
  $a^{*}_i \in \text{min}_{a_i} \text{max}_{a_{-i}} \pi_i(a_i^\$,a_{-i})-\pi_i(a_i,a_{-i}) $.
\end{definition}

@@

% \begin{itemize}
% \item \textcolor{red}{add here:}
%   \begin{itemize}  
%   \item outcome of rationalizability (max-EU)
%   \item outcome of maxmin rule
%   \item outcome of Laplace rule
%   \end{itemize}

@@

Regret minimization is intimately related to the maxmin rule. Let the \emph{negative regret
  transformation} of a strategic game $G$ be the game $G'$ derived from $G$ by replacing the payoff
function $\pi$ of $G$ with $\pi'$ such that $\pi'_i(a_i, a_{-i}) = - reg(a_i,a_{-1})$, i.e.,
the payoff of all players is the pairwise negative regret (under the payoffs from $G$). For
example, the negative regret transformation of the game in Figure~\ref{coordgame1} would have the
following payoff matrix, i.e., the negatives of the pairwise regrets given in
Figure~\ref{coordgame1reg}:

@@

\noindent Acts selected by the maxmin rule for the negative regret transformation of $G$ are
always exactly the acts selected by regret minimization for the original game $G$.\todo{can we
  cite a paper that proves this? should we give a proof ourselves?}

This means that we can think of the maxmin rule and regret minimization as essentially the
same operation on different subjective representations of preference: while maxmin considers
actual objective payoffs (as defined by the modeller's game), regret minimization considers
non-veridical, regret-based subjective preferences. As the example of Figure~\ref{coordgame1}
shows, different subjective preferences can give rise to different choice prescriptions all
else equal. While maxmin is indifferent between $I$ and $II$, regret minimization uniquely
selects $II$.

In sum, a choice rule is a normative prescription of action choices for a rational agent in any
game situation. Choice rules may rest on implicit assumptions about what the DM believes (e.g.,
when interpreting Laplace rule as rational choice under flat beliefs). They may also make
implicit assumptions about what the DM prefers (e.g., when reconstructing regret minimization
in terms of maxmin-choices on the regret transformation). This is rather messy. To disentangle
and avoid confusion, we turn to choice mechanisms.

@@

A \emph{choice mechanism} is a map from choice situations to action choices that, unlike choice
rules, contains an explicit level of subjective representation of a decision situation. For our
purposes, we should think of a choice mechanism as a pair of functions ($\gamma,\kappa$),
where $\gamma$ is a \emph{subjective representation scheme} and $\kappa$ is an \emph{action
  selection function} (see FigureXYZ).\todo{make figure when notation and terminology are
  fixed}\ The subjective representation scheme $\gamma$ maps a game (or more generally any
choice situation) onto a subjective representation of the game (or choice situation) and
other decision-relevant beliefs (e.g., about what the opponent is likely to play). The action
selection function $\kappa$ takes such a subjective representation that $\gamma$ returns and
maps it onto a concrete action choice. The crucial bit for our purposes here is the
representation function $\gamma$, because we are interested in studying which ways of
conceptualizing a decision situation are evolutionarily successful.

A choice mechanism could be a set of distinct heuristics (the DM's \emph{adaptive toolbox})
that have little in common
\citep[e.g.,][]{TverskyKahnemann1981:The-Framing-of-,GigerenzerGoldstein1996:Reasoning-the-F,ScheibehenneRieskamp2013:Testing-the-Ada}. But
to keep matters simple and implementable for a start, we will here only look at choice
mechanisms that differ only with respect to their subjective representation function $\gamma$,
which in turn is conceived of as a \emph{uniform} way of forming preference and belief
\emph{types}: we look at general \emph{subjective representation schemes}.

Formally, given a class $ \mathcal{G} $ of strategic games,
$\gamma_i: \mathcal{G} \rightarrow \Theta \times E$ is a function that assigns a subjective
preference $\theta \in \Theta$ and an epistemic state $e \in E$ to any game
$G \in \mathcal{G}$ for some agent $i$. A subjective preference $\theta$ for game $G$ is just a subjective utility function $\theta:A \rightarrow \mathbb{R}$. Crucially, this subjective utility function can be different from the actual payoff function $\pi^G$ of $G$,
which we interpret as the objective evolutionary fitness. An epistemic state $e$ for game $G$ is some specification of a player's uncertainty about the co-players' behavior. Finally, an
action selection function for agent $i$ is a function
$\kappa_i: \Theta \times E \rightarrow \pow{A_i} $, that associates any subjective
representation of the game $ (\theta,e) \in \Theta \times E$ with a set of prescribed
actions for $i$ in $G$ under $i$'s subjective representation of $G$. We drop the index $i$ for
readability in the following.

We focus on two (\textit{subjective}) \emph{preference types} and two \emph{belief types}. We
view a preference type as a function $\tau: \mathcal{G} \rightarrow \Theta$, and a belief type
as a general disposition to form beliefs about the co-player’s behavior. The two possible
dispositions that we take account of are precise probabilistic belief and imprecise belief, as
we will explain in the next section. Since we are most interested here in studying whether
evolutionary pressure could support non-veridical subjective preferences, the two preference
types that we consider are: the veridical type $\tau^\pi$, who uses the actual payoff function
$\pi$ of the actual game $G$, and the regret type $\tau^{reg}$, who looks at the regret
transformation of the actual game $G$. Finally, the action selection function $\kappa$ that we
use in this paper is \textit{maxmin expected utility} \citep{gilsch89}, for reasons that will be clear
and justified in the following two sections.

@@

% The following subsection describes what these belief types are and how a single action
% selection function (maxmin expected utility) gives rise to an intersting parallel with
% classical choice rules.

@@

Consider the following problem. I have, on the desk in front of me, a bag containing either red or black marbles. There is no further information about the distribution of the two colors. What is then the probability that I will draw a red marble? As in \citet{walley96}, we are not talking about physical probability, but rather about epistemic probability, i.e., the subjective probability that an agent attributes to an event.
In the literature there are two traditional ways of answering this kind of questions. According to standard Bayesianism, one should always be able to determine a precise betting rate (that immediately translates into a precise probability distribution), which is fair, in the sense that the DM is indifferent between betting on or against red at the fair rate, and would bet on red at any rate that is more favourable than the fair rate and against red at any rate that is less favourable. In the case of the bag on my desk, a strict Bayesian, in the absence of any further information, would probably rely on the \textit{principle of insufficient reason} and answer that the DM should have a uniform distribution that considers all the outcomes equipossible. \\
The second answer involves imprecise probabilities and is related to the literature about unmeasurable uncertainty. Many authors argued in favor of this kind of approach \citep[e.g.,][]{levi74,gardsah82,walley96}. Since it is consistent with the information available that all the marbles are red, or that no marble is red, the resulting model should take it into account and specify a lower probability of red $\underline{P}(R)$ and an upper probability of red $\overline{P}(R)$ that define an interval of possible probabilistic beliefs [$\underline{P}(R), \overline{P}(R)$]. Given the information available in this case we would have $\underline{P}(R)=0$ and $\overline{P}(R)=1$. This approach appears particularly relevant in a game theoretical context. Indeed, in a recent paper \citet{BattCerrMM15} write:

@@

In evolutionary game theory, for instance, players in a population obviously face uncertainty about the composition of the population that they are part of, and consequently about the co-player that they are randomly paired with at each round and about the co-players' action. We represent this uncertainty in two different ways that correspond to the two approaches discussed above: a Bayesian uniform distribution and an imprecise probability set. This amounts to having two different belief types, $ \overline{\mu}$ and $ \Delta $, where $\overline{\mu}$ stands for the uniform distribution and $\Delta$ is the probability set, defined as follows:

@@

\item $ \Delta \equiv \Delta(A_{-i}) :=\lbrace \mu \in \mathbb{R}^{A_{-i}}: \sum_{a_{-i} \in A_{-i}} \mu(a_{-i}) = 1 \rbrace$;

\item $\overline{\mu} \in \Delta$ is the probability measure over the co-players' actions such that: $\forall a_{-i} \in A_{-i}$, $\overline{\mu}(a_{-i})= \frac{1}{|A_{-i}|}$.

@@

Hence, we consider two possible epistemic states in the face of uncertainty. The belief type $\Delta$ represents a situation of non-Bayesian radical uncertainty, where the players are not able to narrow down the set of probability distributions to a single one and they consider all of them as possible. The other belief type $\overline{\mu}$ conceptualizes the uncertainty about the population by using the principle of insufficient reason and ascribes equal probability to all the possible alternatives. As we have seen, both options are reasonable and justifiable ways to deal with situations of uncertainty. \\

In decision theory, the imprecise probability model is in line with most of the representation results of decision making under uncertainty \citep[e.g.,][]{gilsch89,KlibMarMuk05,GhirMar02}, and seems justified by empirical observations too.

@@

\begin{fact} \label{fact:singleton probability set}
Whenever the probability set is a singleton, maximinimization of
expected utility coincides with maximization of expected utility.
\end{fact}

@@

By fixing the action selection component to maxmin expected utility, and allowing two belief and
two preference types, we regain all the choice rules considered in sectionNN from a more general and unified standpoint, as it is summed up in
Table~\ref{tab:CMs}. \\
We are now in the position of formally deriving behavior from a pair $(\theta, e) \in \Theta \times E$.

@@

It is then worth stressing that $\kappa(\tau^{\pi}, \Delta)$ turns out to be the actual payoff
maximinimizer, $\kappa(\tau^{reg}, \Delta)$ corresponds to the regret minimizer,
$\kappa(\tau^{\pi}, \overline{\mu})$ is the expected actual payoff maximizer (with flat beliefs), and
$\kappa(\tau^{reg}, \overline{\mu})$ is the expected regret minimizer (with flat beliefs).
Moreover, different subjective preferences and belief types can give rise to identical action predictions,
either in a restricted class of games, or in general. This is also the case here.

\begin{table}[t]
  \centering
  \begin{tabular}{ccc}
    preference type & belief type & corresponding choice rule \\ \midrule
    $\tau^\pi$ & $\overline{\mu}$ & Laplace \\
    $\tau^\pi$ & $\Delta$ & maxmin \\ 
    $\tau^{reg}$ & $\overline{\mu}$ & Laplace \\
    $\tau^{reg}$ & $\Delta$ & regret minimization \\ 
  \end{tabular}
  \caption{Overview relevant choice mechanisms and the corresponding (behaviorally equivalent) classical choice rule}
  \label{tab:CMs}
\end{table}

\begin{fact} \label{fact:maxEU-minReg} For any strategic game $G$ and any arbitrary
  probabilistic belief $\mu \in \Delta$ about the opponents' behavior, the acts
  selected by maximization of expected utility for $G$ under $\mu$ are exactly the acts
  selected by the maximization of expected utility for the negative regret transformation of $G$
  under $\mu$ \citep[e.g.,][]{HalpernPass2012:Iterated-Regret}.
\end{fact}
